---
title: "Homework3"
author: "Matt Futia"
date: "2/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem 1:
Transformed

```{r eval=FALSE}
First String	Second		1.22		3.4
Second			More Text	1.555555	2.2220
Third			x			3			124
```

to

```{r eval=FALSE}
First String,Second,1.22,3.4
Second,More Text,1.555555,2.2220
Third,x,3,124
```

by replacing `\t` with `,`

---

### Problem 2:
Transformed

```{r eval=FALSE}
Ballif, Bryan, University of Vermont
Ellison, Aaron, Harvard Forest
Record, Sydne, Bryn Mawr
```

to

```{r eval=FALSE}
Bryan Ballif (University of Vermont)
Aaron Ellison (Harvard Forest)
Sydne Record (Bryn Mawr)
```

by replacing `(\w+),\s+(\w+),\s+(\w.+)` with `\2 \1 \(\3\)`

---

### Problem 3:
Transformed

```{r eval=FALSE}
0001 Georgia Horseshoe.mp3 0002 Billy In The Lowground.mp3 0003 Cherokee Shuffle.mp3 0004 Walking Cane.mp3
```

to 

```{r eval=FALSE}
0001 Georgia Horseshoe.mp3
0002 Billy In The Lowground.mp3
0003 Cherokee Shuffle.mp3
0004 Walking Cane.mp3
```

by replacing `(.mp3)\s+` with `\1\n`

---

### Problem 4:
Transformed output from problem 3 to 

```{r eval=FALSE}
Georgia Horseshoe_0001.mp3
Billy In The Lowground_0002.mp3
Cherokee Shuffle_0003.mp3
Walking Cane_0004.mp3
```

by replacing `(\d{4})\s(.*)(.mp3)` with `\2_\1\3`

---

### Problem 5:
Transformed

```{r eval=FALSE}
Camponotus,pennsylvanicus,10.2,44
Camponotus,herculeanus,10.5,3
Myrmica,punctiventris,12.2,4
Lasius,neoniger,3.3,55
```

to 

```{r eval=FALSE}
C_pennsylvanicus,44
C_herculeanus,3
M_punctiventris,4
L_neoniger,55
```

by replacing `(\w)\w+,(\w+),\d+.\d(,\d+)` with `\1_\2\3`

---

### Problem 6:
Tranformed the original data set from problem 5 to

```{r eval=FALSE}
C_penn,44
C_herc,3
M_punc,4
L_neon,55
```

by replacing `(\w)\w+,(\w{4})\w+,\d+.\d(,\d+)` with `\1_\2\3`

---

### Problem 7:
Tranformed the original data set from problem 5 to

```{r eval=FALSE}
Campen, 44, 10.2
Camher, 3, 10.5
Myrpun, 4, 12.2
Lasneo, 55, 3.3
```

by replacing `(\w{3})\w+,(\w{3})\w+,(\d+.\d),(\d+)` with `\1\2, \4, \3`
