---
title: 'Homework 7: Data Distributions'
author: "Matt Futia"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Open libraries
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

Generate practice data
```{r}
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

Read in personal data (lake trout length)

```{r}
lt <- read.table("lt_dat.csv", header = T, sep = ",", stringsAsFactors = F)
str(lt)
colnames(lt) <- "myVar"
summary(lt)
```

Histogram of data
```{r}
p1 <- ggplot(data=lt, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

Histogram with empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

Normal probability density
```{r}
# Maximum likelihood parameters for normal
normPars <- fitdistr(lt$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(lt$myVar),len=length(lt$myVar))

stat <- stat_function(aes(x = xval, y = ..y..),
                      fun = dnorm, 
                      color="red",
                      n = length(lt$myVar), 
                      args = list(mean = meanML, sd = sdML))
p1 + stat
```

Exponential probability density
```{r}
expoPars <- fitdistr(lt$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..),
                       fun = dexp, 
                       colour="blue",
                       n = length(lt$myVar), 
                       args = list(rate=rateML))
p1 + stat + stat2
```

Uniform probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), 
                       fun = dunif, colour="darkgreen", 
                       n = length(lt$myVar), 
                       args = list(min=min(lt$myVar), max=max(lt$myVar)))

p1 + stat + stat2 + stat3
```


Gamma probability density
```{r}
gammaPars <- fitdistr(lt$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..),
                       fun = dgamma,
                       color="brown",
                       n = length(lt$myVar),
                       args = list(shape=shapeML, rate=rateML))

p1 + stat + stat2 + stat3 + stat4
```

Beta probability density
```{r}
# make in new plot with rescaled raw data between 0 and 1 for comparison to the beta
pSpecial <- ggplot(data=lt, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

# Maximum likelihood parameters for beta distribution
betaPars <- fitdistr(x=lt$myVar/max(lt$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]; shape1ML
shape2ML <- betaPars$estimate["shape2"]; shape2ML

statSpecial <- stat_function(aes(x = xval, y = ..y..), 
                             fun = dbeta, colour="orchid",
                             n = length(lt$myVar), 
                             args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

Best-fitting distribution: beta distribution

Max likelihood parameters from beta distribution
```{r}
# max likelihood parameters
shape1ML
shape2ML

shp1 <- betaPars$estimate["shape1"]
shp2 <- betaPars$estimate["shape2"]

xval <- seq(0,max(lt$myVar),len=length(lt$myVar))

stat <- stat_function(aes(x = xval, y = ..y..),
                      fun = dnorm, 
                      color="red",
                      n = length(lt$myVar), 
                      args = list(mean = meanML, sd = sdML))
```

Create new dataset with beta maximum likelihood parameters
```{r}
new_dat <- data.frame(rbeta(n = 147, shape1 = shp1, shape2 = shp2))
colnames(new_dat) <- "myVar"

pSpecial2 <- ggplot(data=new_dat, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

# Maximum likelihood parameters for beta distribution
betaPars_2 <- fitdistr(x=new_dat$myVar/max(new_dat$myVar + 0.1),
                       start=list(shape1=1,shape2=2),"beta")
shape1ML_2 <- betaPars_2$estimate["shape1"]; shape1ML_2
shape2ML_2 <- betaPars_2$estimate["shape2"]; shape2ML_2

statSpecial_2 <- stat_function(aes(x = xval, y = ..y..), 
                             fun = dbeta, color="orchid",
                             n = length(new_dat$myVar), 
                             args = list(shape1=shape1ML_2,shape2=shape2ML_2))

par(mfrow = c(2,1))
pSpecial + statSpecial
pSpecial2 + statSpecial_2
```

The histograms from the two datasets appear to have similar shapes based on skewness, but the data from the new dataset are much more evenly spread out than the original data. Based on this, the model appears to overrepresnt many bins while under-representing the most frequent bins from the main model.